---
title: "Agglomerative Hierarchical Clustering"
author: "Sydney Rilum"
date: "3/3/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(naniar)
library(ggdendro)
```

```{r}
# Read in the data
watersheds <- read_csv(here("data", "sbc_lter_registered_stream_chemistry.csv")) %>% 
  replace_with_na_all(condition = ~.x == -999.0)  # replace -999.0 values to NA
```

```{r}
# Create a data frame with a single summary row per site (containing calculated means for each variable)
watersheds_summary <- watersheds %>%
  select(-timestamp_local) %>%  # remove time observations column
  group_by(site_code) %>% 
  summarize_all(mean, na.rm = TRUE)


#### remove sites with Na values????


# Scale the numeric variables data (columns 3:12)
watersheds_scaled <- watersheds_summary %>% 
  select(2:11) %>% 
  scale()
  
# Add rownames (watershed site_code) from watersheds_summary dataset
rownames(watersheds_scaled) <- watersheds_summary$site_code
```

### Perform agglomeratuve hierarchical clustering by complete linkage

```{r}
# Find the Euclidean distances (dissimilarity matrix)
euc_distance <- dist(watersheds_scaled, method = "euclidean")

# Hierarchical clustering by complete linkage
hc_complete <- stats::hclust(euc_distance, method = "complete" )

# Make a dendrogram using ggplot
ggdendrogram(hc_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(x = "Watershed Site Code")
```

**Figure 1.** A dendrogram displaying multivariate clustering for water stream chemistry by Santa Barbara LTER watershed site, created using hierarchical clustering by complete linkage. Water chemistry measurements data set includes dissolved nitrogen (nitrate, ammonium, total dissolved nitrogen), soluble reactive phosphorus (SRP), particulate organic carbon, nitrogen and phosphorus, total suspended sediments, and conductivity measurements from 2000 - 2018. (Data: Santa Barbara Coastal LTER 2019). 

Complete linkage merges clusters by the smallest *maximum* distance between two observations in distinct clusters.

<br>

#### Citation:

[Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative.](https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174)
