---
title: "Parameter Estimation - Wild Fish Catch"
author: "Sydney Rilum"
date: "3/3/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
library(nlstools)
```

```{r}
# Read in the data
fish_catch <- read_csv(here("data", "fish_catch.csv")) %>% 
  row_to_names(row_number = 2) %>%  # convert second row to column headers
  clean_names() %>% 
  slice(-(1:2), -(66:69)) %>%  # remove rows with notes/information
  mutate(years = 0:62) %>%  # Create a new column numbering years starting at 0 (1950 = 0)
  mutate(wild_catch = as.numeric(wild_catch))

```

### Exploratory Graph
```{r}
# Exploratory graph
ggplot(data = fish_catch, aes(x = years, y = wild_catch)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Time (years)", y = "Wild catch (million tons)")

```

**Figure 1.** Exploratory graph of world wild fish catch (in millions of tons) over time (in years) from 1950 - 2012. (Data: ?????????)

In text below the exploratory graph: What type of relationship describes the trend? What does that look like mathematically (include an equation, possibly using LaTeX)? What are your initial estimates for the parameters in the model? 

From the exploratory plot, there appears to be a logistic trend of wild fish catch over time, with an increase in wild catch at a relatively constant rate that eventually plateaus off at around 40 years. 

Mathematically, the wild catch data appears to follow the logistic growth equation: 

$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant


### Finding initial estimates for logistic growth parameters: *K*, *A* and *k*

```{r}
# Estimate the growth constant during exponential phase (to get a starting-point guess for *k*): 
fish_catch_exp <- fish_catch %>% 
  filter(years < 40) %>%  # Get only up to 40 years (exponential growth phase)
  mutate(ln_wild_catch = log(wild_catch))  # log transform the wild_catch data
  
# Model linear to get *k* estimate (the slope of this linear equation is an estimate of the growth rate constant):
lm_k <- lm(ln_wild_catch ~ years, data = fish_catch_exp)
lm_k

# Coefficient (k) ~ 0.04
```

Initial estimates for model parameters:

- $K$ ~ 90
- $A$ ~ 4.23
(90-17.2)/17.2 = 4.23
- $k$ ~ 0.04

Now we have initial estimate for *k* (0.04), and we can estimate *K* ~ 90 and *A* ~ 4.23. These estimates will be used as starting points for interactive algorithms trying to converge on the parameters.


### Using nonlinear least squares to find parameters for logistic model describing wild catch

Nonlinear least squares converges on parameter estimates that minimize the sum of squares of residuals through an iterative algorithm.

```{r}
# Enter model information, with a list of estimated starting parameter values:
fish_catch_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*years)),
              data = fish_catch,
              start = list(K = 90, A = 4.23, r = 0.04),
              trace = TRUE
              )

# See the model summary (null hypothesis: parameter value = 0)
summary(fish_catch_nls)

# Use broom:: functions to get model outputs in tidier format: 
model_output <- broom::tidy(fish_catch_nls)

```

###########Report the parameter outcomes (with units) in text or a table.######################

For Task 2, you will find an equation with parameters estimated by nonlinear least squares for the increase in global wild fish catch from 1950 – 2012. 

The wild catch logistic model with estimated parameters is:
$$P(t) = \frac{100.3}{1+4.32e^{-0.07t}}$$



### Visualize logisitc model over original wild catch data

4) Prepare a finalized (publication quality) graph showing both the original data and your model output.

```{r}
# Make predictions of wild catch for each year in the original data frame
fish_catch_predict <- predict(fish_catch_nls)

# Bind predictions to original data frame
fish_catch_complete <- data.frame(fish_catch, fish_catch_predict)

# Plot original data and predicted values from model together:
ggplot(data = fish_catch_complete, aes(x = years, y = wild_catch)) +
  geom_point() +
  geom_line(aes(x = years, y = fish_catch_predict),
            color = "red") +
  theme_minimal() +
  labs(x = "Time (years)",
       y = "Wild catch (million tons)")

```

**Figure 2.** ????

```{r}
# Find confidence intervals for parameter estimates
fish_catch_ci <- confint2(fish_catch_nls)
fish_catch_ci
```


<br>

#### Citation:

[Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases.](www.fao.org/fishery/topic/16140/en)
