---
title: "Agglomerative Hierarchical Clustering"
author: "Sydney Rilum"
date: "3/3/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(naniar)
library(ggdendro)
```

```{r}
# Read in the data
watersheds <- read_csv(here("data", "sbc_lter_registered_stream_chemistry.csv")) %>% 
  replace_with_na_all(condition = ~.x == -999.0)  # replace -999.0 values with NA
```

### Agglomerative Hierarchical Clustering by Complete Linkage

```{r}
# Create a data frame with a single summary row per site
watersheds_summary <- watersheds %>%
  select(-timestamp_local) %>%  # remove time observations column
  group_by(site_code) %>% 
  summarize_all(mean, na.rm = TRUE) %>%  # calculate means for each variable per site
  drop_na()  # remove sites with NA values

# Scale numeric data
watersheds_scaled <- watersheds_summary %>% 
  select(2:11) %>% 
  scale()
  
# Add rownames (site_code) from watersheds_summary dataset
rownames(watersheds_scaled) <- watersheds_summary$site_code

# Find the Euclidean distances (dissimilarity matrix)
euc_distance <- dist(watersheds_scaled, method = "euclidean")

# Hierarchical clustering by complete linkage
hc_complete <- stats::hclust(euc_distance, method = "complete" )

# Make a dendrogram using ggplot
ggdendrogram(hc_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(x = "Watershed Site Code",
       y = "Euclidean distance of scaled values",
       title = "Multivariate Clustering for Santa Barbara Water Chemistry by Site")
```

**Figure 1.** A dendrogram displaying multivariate clustering for water stream chemistry by Santa Barbara LTER watershed site, created using hierarchical clustering by complete linkage (merging clusters by the smallest maximum distance between two observations in distinct clusters). Water chemistry measurements were taken from 2000 - 2018 for the following variables: dissolved nitrogen (nitrate, ammonium, total dissolved nitrogen), soluble reactive phosphorus, particulate organic carbon, nitrogen and phosphorus, total suspended sediments, and conductivity. (Data: Santa Barbara Coastal LTER 2019). 

<br>

#### Citation:

[Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative.](https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174)
